---
title: Longitudinal LME analysis of the DESCRIBE dataset
author: Francisco J. Martinez-Murcia
date: '\today{}'
output:
  pdf_document:
    toc: 'yes'
    toc_depth: 3
    latex_engine: xelatex
    df_print: tibble
  'bookdown::pdf_document2': default
  'bookdown::html_document2': default
  html_notebook: default
  word_document:
    toc: 'yes'
    toc_depth: '2'
format:
  docx:
    toc: true
    number-sections: true
    highlight-style: github
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
editor_options:
  markdown:
    wrap: 72
---

# TASKS:

-   9 vs 1X
-   each group independently
-   9, 10, 12, 11+13,

9: behavioural variant of FTD 10: PPA not specified 11: non-fluent
variant of PPA 12: logopenic variant of PPA 13: semantic variant of PPA

## Names: 
9 : bvFTD - behavioural variant of FTD 
10: nsPPA - PPA not specified 
11: nfPPA - non-fluent variant of PPA 
12: lvPPA - logopenic variant of PPA 
13: svPPA - semantic variant of PPA

## PCA clusters

-   active_behav: sympathicotonic

-   passive_behav: vagotonic

-   affective: affective symptoms

-   psychotic: psychotic symptoms

# Fitting Linear Mixed Effects to the data: 4 component model.

This is the **LME** analysis of the symptoms of the DESCRIBE dataset,
using groups of symptoms as defined by the PCA rotated model. To do so,
we first import the data. Note that we are only considering the
*sum\_{i}*, with $i=1,...4$ (in the 4-scores model) as response
variables, that correspond respectively to X phenotypes. Also, we are
including just Estimated Years to Onset (EYO), Family, Mutation Type
(Group), Gender and name of the acquisition site (blinded.site):

```{r}
library(kableExtra)
library(dplyr)
options(encoding="UTF-8")
options(huxtable.knitr_output_format='md')
describe_long <- readxl::read_excel('Describe.xlsx')
d <- describe_long[c('Site',	'Visit', 'onset',	'MMSE',	'NPI', 'sum1','sum2','sum3', 'sum4')]

d$Age <- describe_long$`age at visit`
d$Code <- factor(describe_long$Pseudonym)
d$EYO <- describe_long$`years since onset`
d$Education <- describe_long$`years of education`
d$Gender <- factor(describe_long$Sex)
d$Site <- factor(d$Site)
d$BinGroup[describe_long$diagnosis<10] <- "bvFTD"
d$BinGroup[describe_long$diagnosis>=10] <- "PPA" 
d$BinGroup <- factor(d$BinGroup, levels=c('PPA', 'bvFTD'))
d$active_behav <- describe_long$sum1
d$passive_behav <- describe_long$sum2
d$affective <- describe_long$sum3
d$psychotic <- describe_long$sum4


d$Group[describe_long$diagnosis == 9] <- "bvFTD"
d$Group[describe_long$diagnosis == 10] <- "nsPPA"
d$Group[describe_long$diagnosis == 11] <- "nfPPA"
d$Group[describe_long$diagnosis == 12] <- "lvPPA"
d$Group[describe_long$diagnosis == 13] <- "svPPA"
d$Group <- factor(d$Group)

# Clean
d <- d %>%
    group_by(Code) %>%
    mutate(EYO = Age - first(onset)) 
d <- d %>%
    group_by(Code) %>%
    mutate(Education = first(Education)) 
d <- d[!is.na(d$active_behav),]
d <- d[!is.na(d$EYO),] # remove all that have not onset registered.
summary(d)
```

We display the distribution of the factor loadings for each subject over
EYO:

```{r}
library(lattice)
xyplot( active_behav + passive_behav + affective +psychotic  ~ EYO, groups=Group,
       d, aspect= "1.5", layout=c(4,1), type = c("g", "p", "r"),
             index.cond = function(x,y) coef(lm(y ~ x))[1],
             xlab = "EYO",
             ylab = "Sum of Scores")
```

```{r}
library(tidyverse)

data <- d %>%
  group_by(Code) %>%
  filter(any(Visit==1)) %>%
  mutate_at(vars(starts_with("sum")), ~ . - .[Visit == 1]) %>%
  ungroup()

# plot spaghetti plot
ggplot(data, aes(x = Visit, y = active_behav, color = as.factor(Code))) +
    geom_line() +
    scale_color_discrete(name = "Subject") + 
    geom_smooth(se=FALSE, colour="black", size=2) +
    guides(color = FALSE) +
    labs(x = "Visit", y = "active_behav difference from baseline")
```

We first try the simpler model, with only one term depending on EYO, in
which the data is: $$ y = \alpha + \beta_1 EYO + \beta_2 Group + b_1$$
with $b_i = \mu_i$, are the random effects, one intercept per family, as
in the paper.

# Active behavioral phenotype

## Binary - bvFTD vs REST:

We first build and check the baseline model:

```{r Fig1, echo=TRUE, fig.height=10, fig.width=5}
library(lme4)
library(car)
library(performance)
library(jtools)
sc1.bm0 <- lmer(active_behav ~ 1 + EYO*BinGroup + (1|Code), data=d)
summ(sc1.bm0)
model_performance(sc1.bm0)
check_model(sc1.bm0)
```

```{r}
library(interactions)
library(ggplot2)
sc1.bm0 <- lmer(active_behav ~ 1 + EYO + BinGroup + (1|Code), data=d)
summ(sc1.bm0)
qqPlot(resid(sc1.bm0))
Anova(sc1.bm0)
interact_plot(sc1.bm0, pred=EYO, modx=BinGroup, data=d, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r}
sc1.bm1 <- lmer(active_behav ~ 1 + EYO*BinGroup + (1|Code), data=d)
summ(sc1.bm1)
qqPlot(resid(sc1.bm1))
anova(sc1.bm0, sc1.bm1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc1.bm2 <- lmer(active_behav ~ 1 + EYO + BinGroup + Gender + (1|Code), data=d)
summ(sc1.bm2)
qqPlot(resid(sc1.bm2))
model_performance(sc1.bm2)
check_model(sc1.bm2)
```

```{r}
anova(sc1.bm0, sc1.bm2)
Anova(sc1.bm2)
interact_plot(sc1.bm2, pred=EYO, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc1.bm3 <- lmer(active_behav ~ 1 + EYO + BinGroup + Gender*EYO + (1|Code), data=d)
summary(sc1.bm3)
#model_performance(sc1.bm2)
check_model(sc1.bm3)
```

```{r}
anova(sc1.bm2, sc1.bm3)
Anova(sc1.bm3)
interact_plot(sc1.bm3, pred=EYO, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Active Behavioural Phenotype', rug=TRUE, jitter=1)
```


```{r echo=TRUE, fig.height=10, fig.width=5}
sc1.bm4 <- lmer(active_behav ~ 1 + BinGroup*Gender + Gender*EYO + (1|Code), data=d)
summary(sc1.bm4)
#model_performance(sc1.bm2)
check_model(sc1.bm4)
```

```{r}
anova(sc1.bm3, sc1.bm4)
Anova(sc1.bm4)
interact_plot(sc1.bm4, pred=EYO, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Active Behavioural Phenotype', rug=TRUE, jitter=1)
```

Including the random effect of site doesn't change anything.

```{r echo=TRUE, fig.height=10, fig.width=5}
#sc1.bm5 <- lmer(active_behav ~ 1 + BinGroup*Gender + Gender*EYO + (1|Code) +(1|Site), data=d)
sc1.bm5 <- lmer(active_behav ~ 1 + EYO + BinGroup + Gender + (1|Code) +(1|Site), data=d)

summary(sc1.bm5)
#model_performance(sc1.bm2)
check_model(sc1.bm5)
```

```{r}
anova(sc1.bm2, sc1.bm5)
Anova(sc1.bm5)
interact_plot(sc1.bm5, pred=EYO, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Active Behavioural Phenotype', rug=TRUE, jitter=1)

```

Education doesn't have any effect (neither over sc1.bm2 or sc1.bm4
refitted).

```{r echo=TRUE, fig.height=10, fig.width=5}
#sc1.bm5 <- lmer(active_behav ~ 1 + BinGroup*Gender + Gender*EYO + (1|Code) +(1|Site), data=d)
dred.educ <- d[!is.na(d$Education),]
sc1.bm2.refit <- lmer(active_behav ~ 1 + EYO + BinGroup + Gender + (1|Code) , data=dred.educ)
sc1.bm6 <- lmer(active_behav ~ 1 + EYO + BinGroup + Gender + Education + (1|Code) , data=dred.educ)

summary(sc1.bm6)
#model_performance(sc1.bm2)
check_model(sc1.bm6)
```

```{r}

anova(sc1.bm2.refit, sc1.bm6)
Anova(sc1.bm6)
interact_plot(sc1.bm6, pred=EYO, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Active Behavioural Phenotype', rug=TRUE, jitter=1)
```

MMSE have a weak effect (p\<0.1), but it relies on dropping more than
100 patients from the analysis, so we chose not to include MMSE.

```{r echo=TRUE, fig.height=10, fig.width=5}
#sc1.bm5 <- lmer(active_behav ~ 1 + BinGroup*Gender + Gender*EYO + (1|Code) +(1|Site), data=d)
dred.mmse <- d[!is.na(d$MMSE),]
sc1.bm2.refit <- lmer(active_behav ~ 1 +  EYO + BinGroup +Gender + (1|Code) , data=dred.mmse)
sc1.bm7 <- lmer(active_behav ~ 1 +   EYO + BinGroup +Gender + MMSE + (1|Code) , data=dred.mmse)

summary(sc1.bm7)
#model_performance(sc1.bm2)
check_model(sc1.bm7)
```

```{r}

anova(sc1.bm2.refit, sc1.bm7)
Anova(sc1.bm7)
interact_plot(sc1.bm7, pred=MMSE, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Active Behavioural Phenotype', rug=TRUE, jitter=1)
```

Let's see what happens if we use Age instead of EYO.

```{r echo=TRUE, fig.height=10, fig.width=5}

sc1.bm8 <- lmer(active_behav ~ 1 +   Age + BinGroup +Gender + (1|Code) , data=d)

summary(sc1.bm8)
#model_performance(sc1.bm2)
check_model(sc1.bm8)
```

```{r}
anova(sc1.bm2, sc1.bm8)
Anova(sc1.bm8)
interact_plot(sc1.bm8, pred=Age, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=TRUE, y.label="Affective Behavioural Phenotype")

```

```{r}
library(car)
library(jtools)
anova(sc1.bm2, sc1.bm4)
plot_summs(sc1.bm0, sc1.bm1, sc1.bm2, sc1.bm3, sc1.bm4, 
             error_format = "[{conf.low}, {conf.high}]")
export_summs(sc1.bm0, sc1.bm1, sc1.bm2, sc1.bm3, sc1.bm4,  model.names=c('sc1.bm0', 'sc1.bm1', 'sc1.bm2', 'sc1.bm3', 'sc1.bm4'),
             error_format = "[{conf.low}, {conf.high}]")
```

IN order to focus better, just compare fm2 and fm4

```{r}
plot_summs(sc1.bm2, sc1.bm4, 
             error_format = "[{conf.low}, {conf.high}]")
export_summs(sc1.bm2, sc1.bm4,  model.names=c('sc1.bm2', 'sc1.bm4'),
             error_format = "[{conf.low}, {conf.high}]")
```

## Three Grous - bvFTD vs nfPPA+svPPA vs lvPPA:

```{r}
library(dplyr)
dataTer = d[d$Group!="nsPPA",]
dataTer <- dataTer %>%
  mutate(TerGroup = case_when(
    Group == "bvFTD" ~ 'bvFTD',
    Group %in% c('nfPPA', 'svPPA') ~ 'PPA-FTD',
    Group == 'lvPPA' ~ 'lvPPA'
  ))
dataTer$TerGroup <- factor(dataTer$TerGroup)
summary(dataTer)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
library(lme4)
library(car)
library(performance)
sc1.fm0 <- lmer(active_behav ~ 1 + EYO + (1|Code), data=dataTer)
summary(sc1.fm0)
model_performance(sc1.fm0)
check_model(sc1.fm0)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc1.fm1 <- lmer(active_behav ~ 1 + EYO + TerGroup + (1|Code), data=dataTer)
summary(sc1.fm1)
model_performance(sc1.fm1)
anova(sc1.fm0, sc1.fm1)
check_model(sc1.fm1)
```

```{r}

Anova(sc1.fm1)
interact_plot(sc1.fm1, pred=EYO, modx=TerGroup, data=dataTer, interval=TRUE, plot.points=TRUE, y.label='Affecttive Behavioural Phenotype')
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc1.fm2 <- lmer(active_behav ~ 1 + EYO*TerGroup + (1|Code), data=dataTer)
summary(sc1.fm2)
model_performance(sc1.fm2)
anova(sc1.fm1, sc1.fm2)
check_model(sc1.fm2)
```

```{r}
Anova(sc1.fm2)
interact_plot(sc1.fm2, pred=EYO, modx=TerGroup, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, rug.data=dataTer$TerGroup)

```

```{r echo=TRUE, fig.height=10, fig.width=5}

sc1.fm3 <- lmer(active_behav ~ 1 + EYO + TerGroup + Gender + (1|Code), data=dataTer)
summary(sc1.fm3)
model_performance(sc1.fm3)
anova(sc1.fm1, sc1.fm3)
check_model(sc1.fm3)
```

```{r}
Anova(sc1.fm3)
interact_plot(sc1.fm3, pred=EYO, modx=TerGroup, mod2=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
```

Now interactions:

```{r}
sc1.fm4 <- lmer(active_behav ~ 1 + EYO + TerGroup*Gender + (1|Code), data=dataTer)
sc1.fm5 <- lmer(active_behav ~ 1 + EYO*Gender + TerGroup + (1|Code), data=dataTer)
sc1.fm6 <- lmer(active_behav ~ 1 + EYO*Gender + TerGroup*Gender + (1|Code), data=dataTer)
summary(sc1.fm4)
summary(sc1.fm5)
summary(sc1.fm6)
#model_performance(sc1.fm4)
anova(sc1.fm3, sc1.fm4)
anova(sc1.fm3, sc1.fm5)
anova(sc1.fm4, sc1.fm5)
anova(sc1.fm4, sc1.fm6)
anova(sc1.fm5, sc1.fm6)
#check_model(sc1.fm4)
Anova(sc1.fm4)
Anova(sc1.fm5)
Anova(sc1.fm6)
interact_plot(sc1.fm4, pred=EYO, modx=TerGroup, mod2=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
interact_plot(sc1.fm5, pred=EYO, modx=TerGroup, mod2=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
interact_plot(sc1.fm6, pred=EYO, modx=TerGroup, mod2=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
```

LEt's add covariates. is Site a random effect?

```{r echo=TRUE, fig.height=10, fig.width=5}
sc1.fm5 <- lmer(active_behav ~ 1 + EYO + TerGroup*Gender + (1|Code) + (1|Site), data=dataTer)
summary(sc1.fm5)
model_performance(sc1.fm5)
anova(sc1.fm4, sc1.fm5)
check_model(sc1.fm5)
```

```{r}
Anova(sc1.fm5)
interact_plot(sc1.fm5, pred=EYO, modx=TerGroup, mod2=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
```

Site is not a random effect. Let's add education years. Not significant
at all.

```{r echo=TRUE, fig.height=10, fig.width=5}
dataTer.educ <- dataTer[!is.na(dataTer$Education),]
sc1.fm4.refitted <- lmer(active_behav ~ 1 + EYO + TerGroup*Gender + (1|Code), data=dataTer.educ)
sc1.fm6 <- lmer(active_behav ~ 1 + EYO + TerGroup*Gender + Education + (1|Code), data=dataTer.educ)
summary(sc1.fm6)
model_performance(sc1.fm6)
anova(sc1.fm4.refitted, sc1.fm6)
check_model(sc1.fm6)
```

```{r}
Anova(sc1.fm6)
interact_plot(sc1.fm6, pred=EYO, modx=TerGroup, mod2=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
```

Site is not a random effect. Let's add MMSE years. Not significant at
all.

```{r echo=TRUE, fig.height=10, fig.width=5}
dataTer.mmse <- dataTer[!is.na(dataTer$MMSE),]
sc1.fm4.refitted <- lmer(active_behav ~ 1 + EYO + TerGroup*Gender + (1|Code), data=dataTer.mmse)
sc1.fm7 <- lmer(active_behav ~ 1 + EYO + TerGroup*Gender + MMSE + (1|Code), data=dataTer.mmse)
summary(sc1.fm7)
model_performance(sc1.fm7)
anova(sc1.fm4.refitted, sc1.fm7)
check_model(sc1.fm7)
```

```{r}
Anova(sc1.fm7)
interact_plot(sc1.fm7, pred=MMSE, modx=TerGroup, mod2=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r}
library(car)
library(jtools)
anova(sc1.fm2, sc1.fm4)
plot_summs(sc1.fm0, sc1.fm1, sc1.fm2, sc1.fm3, sc1.fm4, 
             error_format = "[{conf.low}, {conf.high}]")
export_summs(sc1.fm0, sc1.fm1, sc1.fm2, sc1.fm3, sc1.fm4,  model.names=c('sc1.fm0', 'sc1.fm1', 'sc1.fm2', 'sc1.fm3', 'sc1.fm4'),
             error_format = "[{conf.low}, {conf.high}]")
```

## Four Groups - bvFTD vs lvPPA vs nfPPA vs svPPA:

```{r}
library(dplyr)
dataQuad = d[d$Group!='nsPPA',]
dataQuad$QuadGroup <- factor(dataQuad$Group)
summary(dataQuad)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
library(lme4)
library(car)
library(performance)
sc1.fm0 <- lmer(active_behav ~ 1 + EYO + (1|Code), data=dataQuad)
summary(sc1.fm0)
model_performance(sc1.fm0)
check_model(sc1.fm0)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc1.fm1 <- lmer(active_behav ~ 1 + EYO + QuadGroup + (1|Code), data=dataQuad)
summary(sc1.fm1)
model_performance(sc1.fm1)
anova(sc1.fm0, sc1.fm1)
check_model(sc1.fm1)
```

```{r}
Anova(sc1.fm1)
interact_plot(sc1.fm1, pred=EYO, modx=QuadGroup, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc1.fm2 <- lmer(active_behav ~ 1 + EYO*QuadGroup + (1|Code), data=dataQuad)
summary(sc1.fm2)
model_performance(sc1.fm2)
anova(sc1.fm1, sc1.fm2)
check_model(sc1.fm2)
```

```{r}
Anova(sc1.fm2)
interact_plot(sc1.fm2, pred=EYO, modx=QuadGroup, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc1.fm3 <- lmer(active_behav ~ 1 + EYO + QuadGroup + Gender + (1|Code), data=dataQuad)
summary(sc1.fm3)
model_performance(sc1.fm3)
anova(sc1.fm1, sc1.fm3)
check_model(sc1.fm3)
```

```{r}
Anova(sc1.fm3)
interact_plot(sc1.fm3, pred=EYO, modx=QuadGroup, mod2=Gender, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
```


```{r}
sc1.fm4 <- lmer(active_behav ~ 1 + EYO + QuadGroup:Gender + QuadGroup + Gender + (1|Code), data=dataQuad)
sc1.fm5 <- lmer(active_behav ~ 1 + EYO*Gender + QuadGroup + (1|Code), data=dataQuad)
sc1.fm6 <- lmer(active_behav ~ 1 + EYO + QuadGroup:Gender + QuadGroup + Gender + EYO:Gender + (1|Code), data=dataQuad)
summary(sc1.fm4)
summary(sc1.fm5)
summary(sc1.fm6)
#model_performance(sc1.fm4)
anova(sc1.fm3, sc1.fm4)
anova(sc1.fm3, sc1.fm5)
anova(sc1.fm4, sc1.fm5)
anova(sc1.fm4, sc1.fm6)
anova(sc1.fm5, sc1.fm6)
#check_model(sc1.fm4)
Anova(sc1.fm4)
Anova(sc1.fm5)
Anova(sc1.fm6)
interact_plot(sc1.fm4, pred=EYO, modx=QuadGroup, mod2=Gender, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
interact_plot(sc1.fm5, pred=EYO, modx=QuadGroup, mod2=Gender, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
interact_plot(sc1.fm6, pred=EYO, modx=QuadGroup, mod2=Gender, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
```

LEt's add covariates. is Site a random effect?

```{r echo=TRUE, fig.height=10, fig.width=5}
sc1.fm5 <- lmer(active_behav ~ 1 + EYO + QuadGroup*Gender + (1|Code) + (1|Site), data=dataQuad)
summary(sc1.fm5)
model_performance(sc1.fm5)
anova(sc1.fm4, sc1.fm5)
check_model(sc1.fm5)
```

```{r}
Anova(sc1.fm5)
interact_plot(sc1.fm5, pred=EYO, modx=QuadGroup, mod2=Gender, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
```

Site is not a random effect. Let's add education years. Not significant
at all.

```{r echo=TRUE, fig.height=10, fig.width=5}
dataQuad.educ <- dataQuad[!is.na(dataQuad$Education),]
sc1.fm6.refitted <- lmer(active_behav ~ 1 + EYO*Gender + QuadGroup*Gender + (1|Code), data=dataQuad.educ)
sc1.fm7 <- lmer(active_behav ~ 1 + EYO*Gender + QuadGroup*Gender + Education + (1|Code), data=dataQuad.educ)
summary(sc1.fm7)
model_performance(sc1.fm7)
anova(sc1.fm6.refitted, sc1.fm7)
check_model(sc1.fm7)
```

```{r}
Anova(sc1.fm7)
interact_plot(sc1.fm7, pred=EYO, modx=QuadGroup, mod2=Gender, data=dataQuad.educ, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
```

Site is not a random effect. Let's add MMSE years. Not significant at
all.

```{r echo=TRUE, fig.height=10, fig.width=5}
dataQuad.mmse <- dataQuad[!is.na(dataQuad$MMSE),]
sc1.fm6.refitted <- lmer(active_behav ~ 1 + EYO*Gender + QuadGroup*Gender + (1|Code), data=dataQuad.mmse)
sc1.fm8 <- lmer(active_behav ~ 1 + EYO*Gender + QuadGroup*Gender + MMSE + (1|Code), data=dataQuad.mmse)
summary(sc1.fm8)
model_performance(sc1.fm8)
anova(sc1.fm6.refitted, sc1.fm8)
check_model(sc1.fm8)
```

```{r}
Anova(sc1.fm8)
interact_plot(sc1.fm8, pred=MMSE, modx=QuadGroup, mod2=Gender, data=dataQuad.mmse, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r}
library(car)
library(jtools)
plot_summs(sc1.fm0, sc1.fm1, sc1.fm2, sc1.fm3, sc1.fm4, sc1.fm5, sc1.fm6, 
             error_format = "[{conf.low}, {conf.high}]")
export_summs(sc1.fm0, sc1.fm1, sc1.fm2, sc1.fm3, sc1.fm4, sc1.fm5, sc1.fm6,  model.names=c('sc1.fm0', 'sc1.fm1', 'sc1.fm2', 'sc1.fm3', 'sc1.fm4', 'sc1.fm5', 'sc1.fm6'),
             error_format = "[{conf.low}, {conf.high}]")
```

To have a clearer view, compare just sc1.fm4 and sc1.fm6.

```{r}
library(car)
library(jtools)
plot_summs(sc1.fm4, sc1.fm6, 
             error_format = "[{conf.low}, {conf.high}]")
export_summs(sc1.fm4, sc1.fm6,  model.names=c('sc1.fm4','sc1.fm6'),
             error_format = "[{conf.low}, {conf.high}]")
Anova(sc1.fm6)
```

# Passive Behavioural Phenotype

## Binary - bvFTD vs REST:

We first build and check the baseline model:

```{r Fig2, echo=TRUE, fig.height=10, fig.width=5}
library(lme4)
library(car)
library(performance)
sc2.bm0 <- lmer(passive_behav ~ 1 + EYO + BinGroup + (1|Code), data=d)
summary(sc2.bm0)
model_performance(sc2.bm0)
check_model(sc2.bm0)
```

```{r}
library(interactions)
library(ggplot2)
sc2.bm0 <- lmer(passive_behav ~ 1 + EYO + BinGroup + (1|Code), data=d)
summary(sc2.bm0)
qqPlot(resid(sc2.bm0))
Anova(sc2.bm0)
interact_plot(sc2.bm0, pred=EYO, modx=BinGroup, data=d, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r}
sc2.bm1 <- lmer(passive_behav ~ 1 + EYO*BinGroup + (1|Code), data=d)
qqPlot(resid(sc2.bm1))
anova(sc2.bm0, sc2.bm1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc2.bm2 <- lmer(passive_behav ~ 1 + EYO + BinGroup + Gender + (1|Code), data=d)
summary(sc2.bm2)
qqPlot(resid(sc2.bm2))
model_performance(sc2.bm2)
check_model(sc2.bm2)
```

```{r}
anova(sc2.bm0, sc2.bm2)
Anova(sc2.bm2)
interact_plot(sc2.bm2, pred=EYO, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc2.bm3 <- lmer(passive_behav ~ 1 + EYO + BinGroup + Gender*EYO + (1|Code), data=d)
summary(sc2.bm3)
#model_performance(sc2.bm2)
check_model(sc2.bm3)
```

```{r}
anova(sc2.bm0, sc2.bm3)
Anova(sc2.bm3)
interact_plot(sc2.bm3, pred=EYO, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc2.bm4 <- lmer(passive_behav ~ 1 + BinGroup*Gender + Gender*EYO + (1|Code), data=d)
summary(sc2.bm4)
#model_performance(sc2.bm2)
check_model(sc2.bm4)
```

```{r}
anova(sc2.bm0, sc2.bm4)
Anova(sc2.bm4)
interact_plot(sc2.bm4, pred=EYO, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

Including the random effect of site doesn't change anything.

```{r echo=TRUE, fig.height=10, fig.width=5}
#sc2.bm5 <- lmer(passive_behav ~ 1 + BinGroup*Gender + Gender*EYO + (1|Code) +(1|Site), data=d)
sc2.bm5 <- lmer(passive_behav ~ 1 + EYO + BinGroup + Gender + (1|Code) +(1|Site), data=d)

summary(sc2.bm5)
#model_performance(sc2.bm2)
check_model(sc2.bm5)
```

```{r}
anova(sc2.bm0, sc2.bm5)
Anova(sc2.bm5)
interact_plot(sc2.bm5, pred=EYO, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

Education doesn't have any effect (neither over sc2.bm0 or sc2.bm0
refitted).

```{r echo=TRUE, fig.height=10, fig.width=5}
#sc2.bm5 <- lmer(passive_behav ~ 1 + BinGroup*Gender + Gender*EYO + (1|Code) +(1|Site), data=d)
dred.educ <- d[!is.na(d$Education),]
sc2.bm0.refit <- lmer(passive_behav ~ 1 + EYO + BinGroup + (1|Code), data=dred.educ)
sc2.bm6 <- lmer(passive_behav ~ 1 + EYO + BinGroup + Education + (1|Code) , data=dred.educ)

summary(sc2.bm6)
#model_performance(sc2.bm2)
check_model(sc2.bm6)
```

```{r}
anova(sc2.bm0.refit, sc2.bm6)
Anova(sc2.bm6)
interact_plot(sc2.bm6, pred=EYO, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

MMSE have a weak effect (p\<0.1), but it relies on dropping more than
100 patients from the analysis, so we chose not to include MMSE.

```{r echo=TRUE, fig.height=10, fig.width=5}
#sc2.bm5 <- lmer(passive_behav ~ 1 + BinGroup*Gender + Gender*EYO + (1|Code) +(1|Site), data=d)
dred.mmse <- d[!is.na(d$MMSE),]
sc2.bm0.refit <- lmer(passive_behav ~ 1 + EYO + BinGroup + (1|Code), data=dred.mmse)
sc2.bm7 <- lmer(passive_behav ~ 1 +   EYO + BinGroup + MMSE + (1|Code) , data=dred.mmse)

summary(sc2.bm7)
#model_performance(sc2.bm2)
check_model(sc2.bm7)
```

```{r}
anova(sc2.bm0.refit, sc2.bm7)
Anova(sc2.bm7)
interact_plot(sc2.bm7, pred=MMSE, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r}
library(car)
library(jtools)
anova(sc2.bm0.refit, sc2.bm7)
plot_summs(sc2.bm0, sc2.bm7, 
             error_format = "[{conf.low}, {conf.high}]")
export_summs(sc2.bm0, sc2.bm1, sc2.bm2, sc2.bm3, sc2.bm7,  model.names=c('sc2.bm0', 'sc2.bm1', 'sc2.bm2', 'sc2.bm3', 'sc2.bm7'),
             error_format = "[{conf.low}, {conf.high}]")
```

```{r}
Anova(sc2.bm0)
```

## Three Grous - bvFTD vs nfPPA+svPPA vs lvPPA:

```{r}
library(dplyr)
dataTer = d[d$Group!="nsPPA",]
dataTer <- dataTer %>%
  mutate(TerGroup = case_when(
    Group == "bvFTD" ~ 'bvFTD',
    Group %in% c('nfPPA', 'svPPA') ~ 'PPA-FTD',
    Group == 'lvPPA' ~ 'lvPPA'
  ))
dataTer$TerGroup <- factor(dataTer$TerGroup)
summary(dataTer)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
library(lme4)
library(car)
library(performance)
sc2.fm0 <- lmer(passive_behav ~ 1 + EYO + (1|Code), data=dataTer)
summary(sc2.fm0)
model_performance(sc2.fm0)
check_model(sc2.fm0)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc2.fm1 <- lmer(passive_behav ~ 1 + EYO + TerGroup + (1|Code), data=dataTer)
summary(sc2.fm1)
model_performance(sc2.fm1)
anova(sc2.fm0, sc2.fm1)
check_model(sc2.fm1)
```

```{r}
Anova(sc2.fm1)
interact_plot(sc2.fm1, pred=EYO, modx=TerGroup, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc2.fm2 <- lmer(passive_behav ~ 1 + EYO*TerGroup + (1|Code), data=dataTer)
summary(sc2.fm2)
model_performance(sc2.fm2)
anova(sc2.fm1, sc2.fm2)
check_model(sc2.fm2)
```

```{r}
Anova(sc2.fm2)
interact_plot(sc2.fm2, pred=EYO, modx=TerGroup, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc2.fm3 <- lmer(passive_behav ~ 1 + EYO + TerGroup + Gender + (1|Code), data=dataTer)
summary(sc2.fm3)
model_performance(sc2.fm3)
anova(sc2.fm1, sc2.fm3)
check_model(sc2.fm3)
```

```{r}
Anova(sc2.fm3)
interact_plot(sc2.fm3, pred=EYO, modx=TerGroup, mod2=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

Now interactions:

```{r}
sc2.fm4 <- lmer(passive_behav ~ 1 + EYO + TerGroup*Gender + (1|Code), data=dataTer)
sc2.fm5 <- lmer(passive_behav ~ 1 + EYO*Gender + TerGroup + (1|Code), data=dataTer)
sc2.fm6 <- lmer(passive_behav ~ 1 + EYO*Gender + TerGroup*Gender + (1|Code), data=dataTer)
summary(sc2.fm4)
summary(sc2.fm5)
summary(sc2.fm6)
#model_performance(sc2.fm4)
anova(sc2.fm1, sc2.fm4)
anova(sc2.fm1, sc2.fm5)
anova(sc2.fm4, sc2.fm5)
anova(sc2.fm4, sc2.fm6)
anova(sc2.fm5, sc2.fm6)
#check_model(sc2.fm4)
Anova(sc2.fm4)
Anova(sc2.fm5)
Anova(sc2.fm6)
interact_plot(sc2.fm4, pred=EYO, modx=TerGroup, mod2=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
interact_plot(sc2.fm5, pred=EYO, modx=TerGroup, mod2=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
interact_plot(sc2.fm6, pred=EYO, modx=TerGroup, mod2=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

LEt's add covariates. is Site a random effect?

```{r echo=TRUE, fig.height=10, fig.width=5}
sc2.fm5 <- lmer(passive_behav ~ 1 + EYO + TerGroup + (1|Code) + (1|Site), data=dataTer)
summary(sc2.fm5)
model_performance(sc2.fm5)
anova(sc2.fm1, sc2.fm5)
check_model(sc2.fm5)
```

```{r}
Anova(sc2.fm5)
interact_plot(sc2.fm5, pred=EYO, modx=TerGroup, mod2=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

Site is not a random effect. Let's add education years. Not significant
at all.

```{r echo=TRUE, fig.height=10, fig.width=5}
dataTer.educ <- dataTer[!is.na(dataTer$Education),]
sc2.fm1.refitted <- lmer(passive_behav ~ 1 + EYO + TerGroup + (1|Code), data=dataTer.educ)
sc2.fm6 <- lmer(passive_behav ~ 1 + EYO + TerGroup + Education + (1|Code), data=dataTer.educ)
summary(sc2.fm6)
model_performance(sc2.fm6)
anova(sc2.fm1.refitted, sc2.fm6)
check_model(sc2.fm6)
```

```{r}
Anova(sc2.fm6)
interact_plot(sc2.fm6, pred=EYO, modx=TerGroup, mod2=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

Site is not a random effect. Let's add MMSE years. Not significant at
all.

```{r echo=TRUE, fig.height=10, fig.width=5}
dataTer.mmse <- dataTer[!is.na(dataTer$MMSE),]
sc2.fm1.refitted <- lmer(passive_behav ~ 1 + EYO + TerGroup + (1|Code), data=dataTer.mmse)
sc2.fm7 <- lmer(passive_behav ~ 1 + EYO + TerGroup + MMSE + (1|Code), data=dataTer.mmse)
summary(sc2.fm7)
model_performance(sc2.fm7)
anova(sc2.fm1.refitted, sc2.fm7)
check_model(sc2.fm7)
```

```{r}
Anova(sc2.fm7)
interact_plot(sc2.fm7, pred=MMSE, modx=TerGroup, mod2=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r}
library(car)
library(jtools)
anova(sc2.fm0, sc2.fm1)
plot_summs(sc2.fm0, sc2.fm1,sc2.fm7,  model.names=c('sc2.fm0', 'sc2.fm1', 'sc2.fm7'),
             error_format = "[{conf.low}, {conf.high}]")
export_summs(sc2.fm0, sc2.fm1, sc2.fm7,  model.names=c('sc2.fm0', 'sc2.fm1', 'sc2.fm7'),
             error_format = "[{conf.low}, {conf.high}]")
```

## Four Groups - bvFTD vs lvPPA vs nfPPA vs svPPA:

```{r}
library(dplyr)
dataQuad = d[d$Group!='nsPPA',]
dataQuad$QuadGroup <- factor(dataQuad$Group)
summary(dataQuad)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
library(lme4)
library(car)
library(performance)
sc2.fm0 <- lmer(passive_behav ~ 1 + EYO + (1|Code), data=dataQuad)
summary(sc2.fm0)
model_performance(sc2.fm0)
check_model(sc2.fm0)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc2.fm1 <- lmer(passive_behav ~ 1 + EYO + QuadGroup + (1|Code), data=dataQuad)
summary(sc2.fm1)
model_performance(sc2.fm1)
anova(sc2.fm0, sc2.fm1)
check_model(sc2.fm1)
```

```{r}
Anova(sc2.fm1)
interact_plot(sc2.fm1, pred=EYO, modx=QuadGroup, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc2.fm2 <- lmer(passive_behav ~ 1 + EYO*QuadGroup + (1|Code), data=dataQuad)
summary(sc2.fm2)
model_performance(sc2.fm2)
anova(sc2.fm1, sc2.fm2)
check_model(sc2.fm2)
```

```{r}
Anova(sc2.fm2)
interact_plot(sc2.fm2, pred=EYO, modx=QuadGroup, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc2.fm3 <- lmer(passive_behav ~ 1 + EYO + QuadGroup + Gender + (1|Code), data=dataQuad)
summary(sc2.fm3)
model_performance(sc2.fm3)
anova(sc2.fm1, sc2.fm3)
check_model(sc2.fm3)
```

```{r}
Anova(sc2.fm3)
interact_plot(sc2.fm3, pred=EYO, modx=QuadGroup, mod2=Gender, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r}
sc2.fm4 <- lmer(passive_behav ~ 1 + EYO + QuadGroup:Gender + QuadGroup + Gender + (1|Code), data=dataQuad)
sc2.fm5 <- lmer(passive_behav ~ 1 + EYO*Gender + QuadGroup + (1|Code), data=dataQuad)
sc2.fm6 <- lmer(passive_behav ~ 1 + EYO + QuadGroup:Gender + QuadGroup + Gender + EYO:Gender + (1|Code), data=dataQuad)
summary(sc2.fm4)
summary(sc2.fm5)
summary(sc2.fm6)
#model_performance(sc2.fm4)
anova(sc2.fm1, sc2.fm4)
anova(sc2.fm1, sc2.fm5)
anova(sc2.fm4, sc2.fm5)
anova(sc2.fm4, sc2.fm6)
anova(sc2.fm5, sc2.fm6)
#check_model(sc2.fm4)
Anova(sc2.fm4)
Anova(sc2.fm5)
Anova(sc2.fm6)
interact_plot(sc2.fm4, pred=EYO, modx=QuadGroup, mod2=Gender, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
interact_plot(sc2.fm5, pred=EYO, modx=QuadGroup, mod2=Gender, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
interact_plot(sc2.fm6, pred=EYO, modx=QuadGroup, mod2=Gender, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

LEt's add covariates. is Site a random effect?

```{r echo=TRUE, fig.height=10, fig.width=5}
sc2.fm5 <- lmer(passive_behav ~ 1 + EYO + QuadGroup + (1|Code) + (1|Site), data=dataQuad)
summary(sc2.fm5)
model_performance(sc2.fm5)
anova(sc2.fm1, sc2.fm5)
check_model(sc2.fm5)
```

```{r}
Anova(sc2.fm5)
interact_plot(sc2.fm5, pred=EYO, modx=QuadGroup, mod2=Gender, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

Site is not a random effect. Let's add education years. Not significant
at all.

```{r echo=TRUE, fig.height=10, fig.width=5}
dataQuad.educ <- dataQuad[!is.na(dataQuad$Education),]
sc2.fm1.refitted <- lmer(passive_behav ~ 1 + EYO + QuadGroup + (1|Code), data=dataQuad.educ)
sc2.fm7 <- lmer(passive_behav ~ 1 + EYO + QuadGroup + Education + (1|Code), data=dataQuad.educ)
summary(sc2.fm7)
model_performance(sc2.fm7)
anova(sc2.fm1.refitted, sc2.fm7)
check_model(sc2.fm7)
```

```{r}
Anova(sc2.fm7)
interact_plot(sc2.fm7, pred=EYO, modx=QuadGroup, mod2=Gender, data=dataQuad.educ, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

Site is not a random effect. Let's add MMSE years. Not significant at
all.

```{r echo=TRUE, fig.height=10, fig.width=5}
dataQuad.mmse <- dataQuad[!is.na(dataQuad$MMSE),]
sc2.fm1.refitted <- lmer(passive_behav ~ 1 + EYO + QuadGroup + (1|Code), data=dataQuad.mmse)
sc2.fm8 <- lmer(passive_behav ~ 1 + EYO + QuadGroup + MMSE + (1|Code), data=dataQuad.mmse)
summary(sc2.fm8)
model_performance(sc2.fm8)
anova(sc2.fm1.refitted, sc2.fm8)
check_model(sc2.fm8)
```

```{r}
Anova(sc2.fm8)
interact_plot(sc2.fm8, pred=MMSE, modx=QuadGroup, mod2=Gender, data=dataQuad.mmse, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
```

```{r}
library(car)
library(jtools)
plot_summs(sc2.fm0, sc2.fm1, sc2.fm8,  model.names=c('sc2.fm0', 'sc2.fm1', 'sc2.fm8'), 
             error_format = "[{conf.low}, {conf.high}]")
export_summs(sc2.fm0, sc2.fm1, sc2.fm8,  model.names=c('sc2.fm0', 'sc2.fm1', 'sc2.fm8'),
             error_format = "[{conf.low}, {conf.high}]")
```

# Affective Phenotype - Affective

## Binary - bvFTD vs REST:

We first build and check the baseline model:

```{r Fig3, echo=TRUE, fig.height=10, fig.width=5}
library(lme4)
library(car)
library(performance)

sc3.bm0 <- lmer(affective ~ 1 + EYO + (1|Code), data=d)
summary(sc3.bm0)
model_performance(sc3.bm0)
check_model(sc3.bm0)
```

```{r}
sc3.bm1 <- lmer(affective ~ 1 + EYO + BinGroup + (1|Code), data=d)
qqPlot(resid(sc3.bm1))
anova(sc3.bm0, sc3.bm1)
summary(sc3.bm1)
Anova(sc3.bm1)
interact_plot(sc3.bm1, pred=EYO, modx=BinGroup, data=d, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)

```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc3.bm2 <- lmer(affective ~ 1 + EYO + Gender + (1|Code), data=d)
summary(sc3.bm2)
qqPlot(resid(sc3.bm2))
model_performance(sc3.bm2)
check_model(sc3.bm2)
```

```{r}
anova(sc3.bm0, sc3.bm2)
Anova(sc3.bm2)
interact_plot(sc3.bm2, pred=EYO, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc3.bm3 <- lmer(affective ~ 1 + EYO + Gender:EYO + (1|Code), data=d)
summary(sc3.bm3)
#model_performance(sc3.bm2)
check_model(sc3.bm3)
```

```{r}
anova(sc3.bm0, sc3.bm3)
Anova(sc3.bm3)
interact_plot(sc3.bm3, pred=EYO, modx=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc3.bm4 <- lmer(affective ~ 1 + BinGroup:Gender + EYO + Gender:EYO + (1|Code), data=d)
summary(sc3.bm4)
#model_performance(sc3.bm2)
check_model(sc3.bm4)
```

```{r}
anova(sc3.bm3, sc3.bm4)
Anova(sc3.bm4)
interact_plot(sc3.bm4, pred=EYO, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

Including the random effect of site doesn't change anything.

```{r echo=TRUE, fig.height=10, fig.width=5}
#sc3.bm5 <- lmer(affective ~ 1 + BinGroup*Gender + Gender*EYO + (1|Code) +(1|Site), data=d)
sc3.bm5 <- lmer(affective ~ 1 + EYO + BinGroup + Gender + (1|Code) +(1|Site), data=d)

summary(sc3.bm5)
#model_performance(sc3.bm2)
check_model(sc3.bm5)
```

```{r}
anova(sc3.bm3, sc3.bm5)
Anova(sc3.bm5)
interact_plot(sc3.bm5, pred=EYO, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

Education doesn't have any effect (neither over sc3.bm0 or sc3.bm0
refitted).

```{r echo=TRUE, fig.height=10, fig.width=5}
#sc3.bm3 <- lmer(affective ~ 1 + EYO + Gender:EYO + (1|Code), data=d)
dred.educ <- d[!is.na(d$Education),]
sc3.bm3.refit <- lmer(affective ~ 1 + EYO + Gender:EYO +  (1|Code), data=dred.educ)
sc3.bm6 <- lmer(affective ~ 1 + EYO + Gender:EYO + Education + (1|Code) , data=dred.educ)

summary(sc3.bm6)
#model_performance(sc3.bm2)
check_model(sc3.bm6)
```

```{r}
anova(sc3.bm3.refit, sc3.bm6)
Anova(sc3.bm6)
interact_plot(sc3.bm6, pred=Education, modx=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

We removed the MMSE, but I want to check if there's a

```{r echo=TRUE, fig.height=10, fig.width=5}
#sc3.bm5 <- lmer(affective ~ 1 + BinGroup*Gender + Gender*EYO + (1|Code) +(1|Site), data=d)
dred.educ <- d[!is.na(d$Education),]
sc3.bm7 <- lmer(affective ~ 1 + EYO + Gender:EYO + Education + Education:Gender + (1|Code), data=dred.educ)

summary(sc3.bm7)
#model_performance(sc3.bm2)
check_model(sc3.bm7)
```

```{r}
anova(sc3.bm6, sc3.bm7)
Anova(sc3.bm7)
interact_plot(sc3.bm7, pred=Education, modx=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

```{r}
library(car)
library(jtools)
anova(sc3.bm3.refit, sc3.bm6)
plot_summs(sc3.bm0, sc3.bm3, sc3.bm6, model.names=c('sc3.bm0', 'sc3.bm3', 'sc3.bm6'),
             error_format = "[{conf.low}, {conf.high}]")
export_summs(sc3.bm0, sc3.bm3, sc3.bm6, model.names=c('sc3.bm0', 'sc3.bm3', 'sc3.bm6'),
             error_format = "[{conf.low}, {conf.high}]")
```

## Three Grous - bvFTD vs nfPPA+svPPA vs lvPPA:

```{r}
library(dplyr)
dataTer = d[d$Group!="nsPPA",]
dataTer <- dataTer %>%
  mutate(TerGroup = case_when(
    Group == "bvFTD" ~ 'bvFTD',
    Group %in% c('nfPPA', 'svPPA') ~ 'PPA-FTD',
    Group == 'lvPPA' ~ 'lvPPA'
  ))
dataTer$TerGroup <- factor(dataTer$TerGroup)
summary(dataTer)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
library(lme4)
library(car)
library(performance)
sc3.fm0 <- lmer(affective ~ 1 + EYO + (1|Code), data=dataTer)
summary(sc3.fm0)
model_performance(sc3.fm0)
check_model(sc3.fm0)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc3.fm1 <- lmer(affective ~ 1 + EYO + TerGroup + (1|Code), data=dataTer)
summary(sc3.fm1)
model_performance(sc3.fm1)
anova(sc3.fm0, sc3.fm1)
check_model(sc3.fm1)
```

```{r}
Anova(sc3.fm1)
interact_plot(sc3.fm1, pred=EYO, modx=TerGroup, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc3.fm2 <- lmer(affective ~ 1 + EYO*TerGroup + (1|Code), data=dataTer)
summary(sc3.fm2)
model_performance(sc3.fm2)
anova(sc3.fm0, sc3.fm2)
check_model(sc3.fm2)
```

```{r}
Anova(sc3.fm2)
interact_plot(sc3.fm2, pred=EYO, modx=TerGroup, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc3.fm3 <- lmer(affective ~ 1 + EYO + Gender + (1|Code), data=dataTer)
summary(sc3.fm3)
model_performance(sc3.fm3)
anova(sc3.fm0, sc3.fm3)
check_model(sc3.fm3)
```

```{r}
Anova(sc3.fm3)
interact_plot(sc3.fm3, pred=EYO, modx=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

Now interactions:

```{r}
sc3.fm4 <- lmer(affective ~ 1 + EYO + EYO:Gender + (1|Code), data=dataTer)
summary(sc3.fm4)
#model_performance(sc3.fm4)
anova(sc3.fm0, sc3.fm4)
#check_model(sc3.fm4)
Anova(sc3.fm4)
interact_plot(sc3.fm4, pred=EYO, modx=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)

```

LEt's add covariates. is Site a random effect?

```{r echo=TRUE, fig.height=10, fig.width=5}
sc3.fm5 <- lmer(affective ~ 1 + EYO + EYO:Gender + (1|Code) + (1|Site), data=dataTer)

summary(sc3.fm5)
model_performance(sc3.fm5)
anova(sc3.fm4, sc3.fm5)
check_model(sc3.fm5)
```

```{r}
Anova(sc3.fm5)
interact_plot(sc3.fm5, pred=EYO, modx=TerGroup, mod2=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

Site is not a random effect. Let's add education years. Not significant
at all.

```{r echo=TRUE, fig.height=10, fig.width=5}
dataTer.educ <- dataTer[!is.na(dataTer$Education),]
sc3.fm4.refitted <- lmer(affective ~1 + EYO + EYO:Gender + (1|Code), data=dataTer.educ)
sc3.fm6 <- lmer(affective ~ 1 + EYO + EYO:Gender + Education + (1|Code), data=dataTer.educ)
summary(sc3.fm6)
model_performance(sc3.fm6)
anova(sc3.fm4.refitted, sc3.fm6)
check_model(sc3.fm6)
```

```{r}
Anova(sc3.fm6)
interact_plot(sc3.fm6, pred=EYO, modx=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

Site is not a random effect. Let's add MMSE years. Not significant at
all.

```{r echo=TRUE, fig.height=10, fig.width=5}
sc3.fm7 <- lmer(affective ~  1 + EYO + EYO:Gender + Education + Education:Gender + (1|Code), data=dataTer.educ)

summary(sc3.fm7)
model_performance(sc3.fm7)
anova(sc3.fm6, sc3.fm7)
check_model(sc3.fm7)
```

```{r}
Anova(sc3.fm7)
interact_plot(sc3.fm7, pred=Education, modx=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

```{r}
library(car)
library(jtools)
anova(sc3.fm0, sc3.fm1)
plot_summs(sc3.fm0, sc3.fm4,sc3.fm6,  model.names=c('sc3.fm0', 'sc3.fm4', 'sc3.fm6'),
             error_format = "[{conf.low}, {conf.high}]")
export_summs(sc3.fm0, sc3.fm4, sc3.fm6,  model.names=c('sc3.fm0', 'sc3.fm4', 'sc3.fm6'),
             error_format = "[{conf.low}, {conf.high}]")
```

## Four Groups - bvFTD vs lvPPA vs nfPPA vs svPPA:

```{r}
library(dplyr)
dataQuad = d[d$Group!='nsPPA',]
dataQuad$QuadGroup <- factor(dataQuad$Group)
summary(dataQuad)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
library(lme4)
library(car)
library(performance)
sc3.fm0 <- lmer(affective ~ 1 + EYO + (1|Code), data=dataQuad)
summary(sc3.fm0)
model_performance(sc3.fm0)
check_model(sc3.fm0)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc3.fm1 <- lmer(affective ~ 1 + EYO + QuadGroup + (1|Code), data=dataQuad)
summary(sc3.fm1)
model_performance(sc3.fm1)
anova(sc3.fm0, sc3.fm1)
check_model(sc3.fm1)
```

```{r}
Anova(sc3.fm1)
interact_plot(sc3.fm1, pred=EYO, modx=QuadGroup, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc3.fm2 <- lmer(affective ~ 1 + EYO*QuadGroup + (1|Code), data=dataQuad)
summary(sc3.fm2)
model_performance(sc3.fm2)
anova(sc3.fm0, sc3.fm2)
check_model(sc3.fm2)
```

```{r}
Anova(sc3.fm2)
interact_plot(sc3.fm2, pred=EYO, modx=QuadGroup, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc3.fm3 <- lmer(affective ~ 1 + EYO + Gender + (1|Code), data=dataQuad)
summary(sc3.fm3)
model_performance(sc3.fm3)
anova(sc3.fm0, sc3.fm3)
check_model(sc3.fm3)
```

```{r}
Anova(sc3.fm3)
interact_plot(sc3.fm3, pred=EYO, modx=Gender, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

```{r}
sc3.fm4 <- lmer(affective ~ 1 + EYO + EYO:Gender + (1|Code), data=dataQuad)
summary(sc3.fm4)
#model_performance(sc3.fm4)
anova(sc3.fm0, sc3.fm4)
#check_model(sc3.fm4)
Anova(sc3.fm4)
interact_plot(sc3.fm4, pred=EYO, modx=Gender, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

FROM THIS MOMENT ON IT IS THE SAME AS THE PREVIOUS ANALYSIS.

# Psychotic Phenotype - Psychotic

## Binary - bvFTD vs REST:

We first build and check the baseline model:

```{r Fig4, echo=TRUE, fig.height=10, fig.width=5}
library(lme4)
library(car)
library(performance)
d.symp = d[d$psychotic>0,]
sc4.bm0 <- lmer(psychotic ~ 1 + EYO + (1|Code), data=d.symp)
summary(sc4.bm0)
Anova(sc4.bm0)
check_model(sc4.bm0)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
library(interactions)
library(ggplot2)
sc4.bm1 <- lmer(psychotic ~ 1 + EYO + BinGroup + (1|Code), data=d.symp)
summary(sc4.bm1)
qqPlot(resid(sc4.bm1))
Anova(sc4.bm1)
check_model(sc4.bm1)
```

```{r}
interact_plot(sc4.bm1, pred=EYO, modx=BinGroup, data=d.symp, interval=TRUE, plot.points=FALSE, y.label='Psychotic Phenotype', rug=TRUE, jitter=1)
```

```{r}
sc4.bm2 <- lmer(psychotic ~ 1 + EYO*BinGroup + (1|Code), data=d.symp)
qqPlot(resid(sc4.bm2))
anova(sc4.bm1, sc4.bm2)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc4.bm3 <- lmer(psychotic ~ 1 + EYO + Gender + (1|Code), data=d.symp)
summary(sc4.bm3)
qqPlot(resid(sc4.bm3))
model_performance(sc4.bm3)
check_model(sc4.bm3)
```

```{r}
anova(sc4.bm0, sc4.bm3)
Anova(sc4.bm3)
interact_plot(sc4.bm3, pred=EYO, modx=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Psychotic Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc4.bm4 <- lmer(psychotic ~ 1 + EYO + Gender:EYO + (1|Code), data=d.symp)
summary(sc4.bm4)
#model_performance(sc4.bm2)
check_model(sc4.bm4)
```

```{r}
anova(sc4.bm0, sc4.bm4)
Anova(sc4.bm4)
interact_plot(sc4.bm4, pred=EYO, modx=Gender, data=d.symp, interval=TRUE, plot.points=FALSE, y.label='Psychotic Phenotype', rug=TRUE, jitter=1)
```

Education doesn't have any effect (neither over sc4.bm0 or sc4.bm0
refitted).

```{r echo=TRUE, fig.height=10, fig.width=5}
#sc4.bm5 <- lmer(psychotic ~ 1 + BinGroup*Gender + Gender*EYO + (1|Code) +(1|Site), data=d)
dred.educ <- d.symp[!is.na(d.symp$Education),]
sc4.bm0.refit <- lmer(psychotic ~ 1 + EYO + (1|Code), data=dred.educ)
sc4.bm6 <- lmer(psychotic ~ 1 + EYO + Education + (1|Code) , data=dred.educ)

summary(sc4.bm6)
#model_performance(sc4.bm2)
check_model(sc4.bm6)
```

```{r}
anova(sc4.bm0.refit, sc4.bm6)
Anova(sc4.bm6)
interact_plot(sc4.bm6, pred=Education, modx=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Psychotic Phenotype', rug=TRUE, jitter=1)
```

```{r}
library(car)
library(jtools)
anova(sc4.bm0, sc4.bm1)
plot_summs(sc4.bm0, sc4.bm1,  model.names=c('sc4.bm0', 'sc4.bm1'), 
             error_format = "[{conf.low}, {conf.high}]")
export_summs(sc4.bm0, sc4.bm1,  model.names=c('sc4.bm0', 'sc4.bm1'),
             error_format = "[{conf.low}, {conf.high}]")
```

## Three Grous - bvFTD vs nfPPA+svPPA vs lvPPA:

```{r}
library(dplyr)
dataTer = d[d$Group!="nsPPA",]
dataTer = dataTer[dataTer$psychotic>0,]
dataTer <- dataTer %>%
  mutate(TerGroup = case_when(
    Group == "bvFTD" ~ 'bvFTD',
    Group %in% c('nfPPA', 'svPPA') ~ 'PPA-FTD',
    Group == 'lvPPA' ~ 'lvPPA'
  ))
dataTer$TerGroup <- factor(dataTer$TerGroup)
summary(dataTer)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
library(lme4)
library(car)
library(performance)
sc4.fm0 <- lmer(psychotic ~ 1 + EYO + (1|Code), data=dataTer)
summary(sc4.fm0)
model_performance(sc4.fm0)
check_model(sc4.fm0)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc4.fm1 <- lmer(psychotic ~ 1 + EYO + TerGroup + (1|Code), data=dataTer)
summary(sc4.fm1)
model_performance(sc4.fm1)
anova(sc4.fm0, sc4.fm1)
check_model(sc4.fm1)
```

```{r}
Anova(sc4.fm1)
interact_plot(sc4.fm1, pred=EYO, modx=TerGroup, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Psychotic Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc4.fm2 <- lmer(psychotic ~ 1 + EYO*TerGroup + (1|Code), data=dataTer)
summary(sc4.fm2)
model_performance(sc4.fm2)
anova(sc4.fm1, sc4.fm2)
check_model(sc4.fm2)
```

```{r}
Anova(sc4.fm2)
interact_plot(sc4.fm2, pred=EYO, modx=TerGroup, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Psychotic Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc4.fm3 <- lmer(psychotic ~ 1 + EYO + Gender + (1|Code), data=dataTer)
summary(sc4.fm3)
model_performance(sc4.fm3)
anova(sc4.fm0, sc4.fm3)
check_model(sc4.fm3)
```

```{r}
Anova(sc4.fm3)
interact_plot(sc4.fm3, pred=EYO, modx=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Psychotic Phenotype', rug=TRUE, jitter=1)
```

Site is not a random effect. Let's add education years. Not significant
at all.

```{r echo=TRUE, fig.height=10, fig.width=5}
dataTer.educ <- dataTer[!is.na(dataTer$Education),]
sc4.fm0.refitted <- lmer(psychotic ~ 1 + EYO + (1|Code), data=dataTer.educ)
sc4.fm6 <- lmer(psychotic ~ 1 + EYO + Education + (1|Code), data=dataTer.educ)
summary(sc4.fm6)
model_performance(sc4.fm6)
anova(sc4.fm0.refitted, sc4.fm6)
check_model(sc4.fm6)
```

```{r}
Anova(sc4.fm6)
interact_plot(sc4.fm6, pred=EYO, modx=Gender, data=dataTer, interval=TRUE, plot.points=FALSE, y.label='Psychotic Phenotype', rug=TRUE, jitter=1)
```

## Four Groups - bvFTD vs lvPPA vs nfPPA vs svPPA:

```{r}
library(dplyr)
dataQuad = d[d$Group!='nsPPA',]
dataQuad = dataQuad[dataQuad$psychotic>0,]
dataQuad$QuadGroup <- factor(dataQuad$Group)
summary(dataQuad)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
library(lme4)
library(car)
library(performance)
sc4.fm0 <- lmer(psychotic ~ 1 + EYO + (1|Code), data=dataQuad)
summary(sc4.fm0)
model_performance(sc4.fm0)
check_model(sc4.fm0)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc4.fm1 <- lmer(psychotic ~ 1 + EYO + QuadGroup + (1|Code), data=dataQuad)
summary(sc4.fm1)
model_performance(sc4.fm1)
anova(sc4.fm0, sc4.fm1)
check_model(sc4.fm1)
```

```{r}
Anova(sc4.fm1)
interact_plot(sc4.fm1, pred=EYO, modx=QuadGroup, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Psychotic Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc4.fm2 <- lmer(psychotic ~ 1 + EYO*QuadGroup + (1|Code), data=dataQuad)
summary(sc4.fm2)
model_performance(sc4.fm2)
anova(sc4.fm1, sc4.fm2)
check_model(sc4.fm2)
```

```{r}
Anova(sc4.fm2)
interact_plot(sc4.fm2, pred=EYO, modx=QuadGroup, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Psychotic Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
sc4.fm3 <- lmer(psychotic ~ 1 + EYO + Gender + (1|Code), data=dataQuad)
summary(sc4.fm3)
model_performance(sc4.fm3)
anova(sc4.fm0, sc4.fm3)
check_model(sc4.fm3)
```

```{r}
Anova(sc4.fm3)
interact_plot(sc4.fm3, pred=EYO, modx=Gender, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Psychotic Phenotype', rug=TRUE, jitter=1)
```

```{r echo=TRUE, fig.height=10, fig.width=5}
dataQuad.educ <- dataQuad[!is.na(dataQuad$Education),]
sc4.fm1.refitted <- lmer(psychotic ~ 1 + EYO  + (1|Code), data=dataQuad.educ)
sc4.fm7 <- lmer(psychotic ~ 1 + EYO + Education + (1|Code), data=dataQuad.educ)
summary(sc4.fm7)
model_performance(sc4.fm7)
anova(sc4.fm1.refitted, sc4.fm7)
check_model(sc4.fm7)
```

```{r}
Anova(sc4.fm7)
interact_plot(sc4.fm7, pred=Education, modx=Gender, data=dataQuad.educ, interval=TRUE, plot.points=FALSE, y.label='Psychotic Phenotype', rug=TRUE, jitter=1)
```

```{r}
library(car)
library(jtools)
plot_summs(sc4.fm0, sc4.fm1, sc4.fm3,  model.names=c('sc4.fm0', 'sc4.fm1', 'sc4.fm3'), 
             error_format = "[{conf.low}, {conf.high}]")
export_summs(sc4.fm0, sc4.fm1, sc4.fm3,  model.names=c('sc4.fm0', 'sc4.fm2', 'sc4.fm3'),
             error_format = "[{conf.low}, {conf.high}]")
```

# Definitive Models

```{r}
library(lme4)
library(car)
library(performance)
library(car)
library(jtools)
# Refit data
# Data ternary
dataTer = d[d$Group!="nsPPA",]
dataTer$original_group <- dataTer$Group
dataTer <- dataTer %>%
  mutate(TerGroup = case_when(
    Group == "bvFTD" ~ 'bvFTD',
    Group %in% c('nfPPA', 'svPPA') ~ 'PPA-FTD',
    Group == 'lvPPA' ~ 'lvPPA'
  ))
dataTer$TerGroup <- factor(dataTer$TerGroup)
dataTer$Group <- factor(dataTer$TerGroup)
dataTer.educ <- dataTer[!is.na(dataTer$Education),]

# Data quaternary 
dataQuad = d[d$Group!='nsPPA',]
dataQuad$original_group <- dataQuad$Group
dataQuad$QuadGroup <- factor(dataQuad$Group)
dataQuad$Group <- factor(dataQuad$QuadGroup)
```

```{r}
sc1.fm4 <- lmer(active_behav ~ 1 + EYO + Group:Gender + Group + Gender + (1|Code), data=dataQuad)
summ(sc1.fm4)
```

```{r}
sc2.fm1 <- lmer(passive_behav ~ 1 + EYO + Group + (1|Code), data=dataQuad)
summ(sc2.fm1)
```

```{r}
sc3.fm6 <- lmer(affective ~ 1 + EYO + EYO:Gender + Education + (1|Code), data=dataTer.educ)
summ(sc3.fm6)
```

```{r}
export_summs(sc1.fm4, sc2.fm1, sc3.fm6, model.names=c('sc1.fm4', 'sc2.fm1', 'sc3.fm6'),
             error_format = "p={p.value} [{conf.low}, {conf.high}]", number_format = 3)
```

## Figures

```{r}
interact_plot(sc1.fm4, pred=EYO, modx=Group, mod2=Gender, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Affecttive Behavioural Phenotype', rug=TRUE, jitter=1)
interact_plot(sc2.fm1, pred=EYO, modx=Group, data=dataQuad, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
interact_plot(sc3.fm6, pred=EYO, modx=Gender, data=dataTer.educ, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```

## N-way Significance testing with Monte Carlo simulations

A N-way empirical significance was estimated from a Monte-Carlo
sampling of the models for each sum score every 1 years in the EYO range
from -0 to 16 in order to identify each signs degree of
differentiation. First we load the packages used for sampling:

And we perform the following: Do the three-way empirical p-value:

$$P(A-B)>0$$

where A and B are:

$$A = dens(SIM\_FAC_i)$$


the probability mass of the sampling from the LME model.

```{r}
library(effects)

plot_effect <- function(model, title, transform, main.effect="Group", ylims=c(-1, 15), eyo_basis=seq(from=0, to=15, by=1),
                        co.effect=NULL,
                        colors=c("bvFTD" = "#49b7fc", "lvPPA" = "#f38000", "nfPPA"="#59d796", "svPPA"="#ff0083")){
  effects <- c(main.effect, "EYO")
  if (!is.null(co.effect)) {
    # Concatenate co.effect to the effects list
    effects <- c(effects, co.effect)
  }
  ee<-Effect(effects, model, partial.residuals=T, xlevels=list(EYO=eyo_basis))
  theme_set(theme_light())
  eetac <- as.data.frame(ee)
  levels.d <- levels(factor(unique(eetac[[main.effect]])))
  eetac[c("fit", "se", "lower","upper")] <- transform(eetac[c("fit", "se", "lower","upper")])
  jarl <- matrix(eetac[['lower']], length(eetac[['lower']])/length(levels.d), length(levels.d), byrow=T)
  jarl.df = list()
  for (i in 1:length(levels.d)){
    jarl.df[[levels.d[i]]] <- jarl[,i]
  }
  #jarl.df <- as.data.frame(jarl.df, col.names=names(jarl.df))<0
  #jarl.df <- cbind(unique(eetac['EYO']),jarl.df)
  #max.bvFTD <- tail(jarl.df[max(which(jarl.df$bvFTD > 0)), 'EYO'], 1)
  #max.lvPPA <- tail(jarl.df[max(which(jarl.df$lvPPA > 0)), 'EYO'], 1)
  #max.nfPPA <- tail(jarl.df[max(which(jarl.df$nfPPA > 0)), 'EYO'], 1)
  #max.svPPA <- tail(jarl.df[max(which(jarl.df$svPPA > 0)), 'EYO'], 1)
  tacmpfig <- ggplot(eetac,
          aes_string("EYO", "fit", colour=main.effect, fill=main.effect))+
      geom_line(size=1)+
       ## colour=NA suppresses edges of the ribbon
      geom_ribbon(colour=NA,alpha=0.1,
                              aes(ymin=lower,ymax=upper))+
       ## add rug plot based on original data
       geom_rug(aes_string(x="EYO",y=ee$response), data=ee$data, sides="b", position='jitter')+ 
    coord_cartesian(ylim=ylims)+labs(title=title, y='Sum score magnitude', x='Estimated Years to Symptom Onset (EYO)')+
    #geom_vline(aes(xintercept = max.bvFTD), color = "#49b7fc", linetype='dashed')+
    #geom_vline(aes(xintercept = max.nfPPA), color = "#59d796", linetype='dashed')+ 
    #geom_vline(aes(xintercept = max.lvPPA), color = "#f38000", linetype='dashed')+ 
    #geom_vline(aes(xintercept = max.svPPA), color = "#ff0083", linetype='dashed')+
    scale_colour_manual(values = colors)+
    scale_fill_manual(values = colors)
  if (!is.null(co.effect)) {
    # Add faceting based on the co.effect variable
    tacmpfig <- tacmpfig + facet_wrap(as.formula(paste("~", co.effect)))
  }
  return(tacmpfig)
}
```


```{r}
get_preds <- function(model, data, transform, main.effect="Group", eyo_basis=seq(from=0, to=15, by=1), N=1e4 ) {
  eyo_seq = rep(eyo_basis, N)
  totsample <- length(eyo_basis)*N # cuidao
  levels = levels(data[[main.effect]])
  Ngroups = length(levels)
  ncols=length(eyo_basis)
  # co-variates
  #Gender = factor(rep("m", N), levels=levels(data$Gender))
  if (main.effect=='Group'){
    Gender = sample(data$Gender, N, replace=TRUE)
  }else if (main.effect=='Gender'){
    Group = sample(data$Group, N, replace=TRUE)
  }
  code = sample(data$Code, N, replace=TRUE)
  bs = sample(sort(data$Site)[1:50], N, replace=TRUE)
  edu = sample(data$Education, N, replace=TRUE)
  
  results = list()
  for (i in 1:Ngroups) {
    newdata <- data.frame(EYO=eyo_seq, Code=code, Site=bs, Education=edu)
    newdata[[main.effect]] = factor(rep(levels[i], N), levels=levels)
    if (main.effect=='Group'){
      newdata$Gender = sample(data$Gender, N, replace=TRUE)
    }else if (main.effect=='Gender'){
      newdata$Group = sample(data$Group, N, replace=TRUE)
    }
    regular <- predict(model, newdata=newdata, re.form=(~0), allow.new.levels=TRUE)
    results[[levels[i]]] <- matrix(transform(regular), N, ncols, byrow=T)
  }
  
  return(results)
}

unlog <- function(x){
  return(10^x -1)
}

iident <- function(x){
  return(x)
}

plot_preds <- function(preds, eyo_basis=seq(from=-40, to=20, by=5), N=1e4, ylim=c(-5,15), title='', colors=c('red', 'blue', 'green', 'violet'), PIrange=.95){
  ncols=length(eyo_basis)
  levels <- names(preds)
  eyo_seq <- rep(eyo_basis, N)
  eyo_matrix_form <- matrix(eyo_seq, N, ncols, byrow=TRUE)
  plot(NULL, xlim=c(min(eyo_basis), max(eyo_basis)), ylim=ylim, ylab=title, xlab='EYO')
  for (i in 1:length(preds)) {
    res <- preds[[levels[i]]]
    preds_mu <- apply(res, 2, mean)
    lines(eyo_matrix_form[1,], preds_mu, col=colors[i])
    preds_PI <- apply(rescaler.data.frame, 2, PI, PIrange)
    shade(preds_PI, eyo_matrix_form[1,], col=col.alpha(colors[i], .2))
    preds_PI_66 <- apply(res, 2, PI, .66)
    shade(preds_PI_66, eyo_matrix_form[1,], col=col.alpha(colors[i], .2))
  }
  legend("topleft", legend = levels,
               text.width = strwidth("1,000,000"),
               title = "Groups")
}


nwayttest <- function(preds, N=1e4){
  final <- data.frame(EYO=eyo_basis)
  levels <- names(preds)
  combinations <- t(combn(levels,2))
  for (i in 1:length(combinations[,1])) {
    com <- combinations[i,]
    comName <- paste(com[1], '.vs.', com[2], sep='')
    dd <- c()
    for (col in 1:length(eyo_basis)){
      sd.1 <- max(sd(preds[[com[1]]][,col]), 0.05)
      sd.2 <- max(sd(preds[[com[2]]][,col]), 0.05)
      d.1 <- rnorm(N, mean(preds[[com[1]]][,col]), sd.1)
      d.2 <- rnorm(N, mean(preds[[com[2]]][,col]), sd.2)
      val <- min(sum(d.1<d.2), sum(d.1>d.2))
      dd <- c(dd, (val+1)/(N+1))
    }
    final[[comName]] <- round(dd*N)/N
  }
  return(final)
}
```

### Active Behavioural Phenotype

```{r}
eyo_basis <- seq(from=0, to=15, by=1)
# N=1e4
# res <- get_preds(sc1.fm4, dataQuad, iident, eyo_basis=eyo_basis, N=N)
# kbl(nwayttest(res), booktabs = T, caption="p-values (four-way hypothesis testing) for the pairwise comparison between groups at each EYO (LMM - Active Behavioural phenotype)") %>%
# kable_styling(latex_options = c("striped", "scale_down"))
# 
# res2 <- get_preds(sc1.fm4, dataQuad, iident, eyo_basis=eyo_basis, main.effect="Gender", N=N)
# kbl(nwayttest(res2), booktabs = T, caption="p-values (four-way hypothesis testing) for the pairwise comparison between Gender at each EYO (LMM - Active Behavioural phenotype)") %>%
# kable_styling(latex_options = c("striped"))

plot_effect(sc1.fm4, transform=iident, main.effect="Group", co.effect="Gender", title='Active Behavioural phenotype', ylims=c(-1, 5), eyo_basis=eyo_basis)
```
With emmeans: 

```{r}
library(emmeans)
library(dplyr)
library(tidyr)
emmip(sc1.fm4, Group ~ Gender)
Anova(sc1.fm4)

#EMM = emmeans(sc1.fm4,  pairwise ~ Gender*Group | EYO)#, at=list(EYO=seq(from=0, to=15, by=1)))
# xprt = print(pairs(EMM), export = TRUE)
# create a pairwise contrast and export to variable in data.frame form: 
EMM <- emmeans(sc1.fm4, specs=c('Group', 'EYO'), at=list(EYO=eyo_basis))
xprt <- as.data.frame(contrast(EMM, method='pairwise', simple='Group'))
kbl(xprt %>% pivot_wider(id_cols=EYO, names_from=c('contrast'), values_from='p.value'), booktabs = T, caption="p-values (four-way hypothesis testing) for the pairwise comparison between groups at each EYO (LMM - Active Behavioural phenotype)")%>%
kable_styling(latex_options = c("striped", "scale_down"))
#xprt <- print(contrast(EMM, method='pairwise', simple='Group'), export=TRUE)


EMM <- emmeans(sc1.fm4, specs=c('Group', 'Gender', 'EYO'), at=list(EYO=eyo_basis))
xprt <- as.data.frame(contrast(EMM, method='pairwise', simple='Group'))
kbl(xprt %>% pivot_wider(id_cols=EYO, names_from=c('contrast', 'Gender'), values_from='p.value'), booktabs = T, caption="p-values (four-way hypothesis testing) for the pairwise comparison between combinations of Group/Gender at each EYO (LMM - Active Behavioural phenotype)")%>%
kable_styling(latex_options = c("striped", "scale_down"))
#xprt <- summary(EMM)
#t(xprt$contrasts %>% pivot_wider(names_from=EYO, values_from='p.value'))
#kbl(t(xprt$contrasts %>% pivot_wider(names_from=EYO, values_from=c("p.value"))))

#kbl(xprt$summary, booktabs = T, caption="p-values (four-way hypothesis testing) for the pairwise comparison between groups at each EYO (LMM - Active Behavioural phenotype)") %>%
#kable_styling(latex_options = c("striped", "scale_down"))
#emmip(sc1.fm4, Group ~ EYO | Gender, mult.name = "Group", cov.reduce = FALSE, CIs=TRUE)

```
#### Contrasts
```{r}
# sc1.fm4 <- lmer(active_behav ~ 1 + EYO + Group:Gender + Group + Gender + (1|Code), data=dataQuad)
joint_tests(sc1.fm4)
emmip(sc1.fm4, Group ~ Gender)
```

```{r}

EMM <- emmeans(sc1.fm4, ~ factor(Group)*factor(Gender), at=list(EYO=eyo_basis))
pairs(EMM, simple = "Group")    # compare treats for each dose -- "simple effects"
pairs(EMM, simple = "Gender")     # compare doses for each treat
eff_size(EMM, sigma = sigma(sc1.fm4), edf = df.residual(sc1.fm4))
plot(EMM, comparisons = TRUE)

```

Degrees-of-freedom method: kenward-roger 
P value adjustment: tukey method for comparing a family of 4 estimates 

### Passive Behavioural Phenotype
```{r}
eyo_basis <- seq(from=0, to=15, by=1)
#N=1e4
#res <- get_preds(sc2.fm1, dataQuad, iident,  eyo_basis=eyo_basis, N=N)
#kbl(nwayttest(res), booktabs = T, caption="p-values (four-way hypothesis testing) for the pairwise comparison between groups at each EYO (LMM - Passive Behavioural phenotype)") %>%
#kable_styling(latex_options = c("striped", "scale_down"))
plot_effect(sc2.fm1, transform=iident, title='Passive Behavioural phenotype', ylims=c(-1, 5), eyo_basis=eyo_basis)
```
#### Contrasts
```{r}
# sc2.fm1 <- lmer(passive_behav ~ 1 + EYO + Group + (1|Code), data=dataQuad)
joint_tests(sc2.fm1)
emmip(sc2.fm1, Group ~ EYO, at=list(EYO=eyo_basis))
```

```{r}

EMM <- emmeans(sc2.fm1, ~ factor(Group) + EYO)
pairs(EMM, simple = "Group")    # compare treats for each dose -- "simple effects"
pairs(EMM, simple = "EYO")     # compare doses for each treat
eff_size(EMM, sigma = sigma(sc1.fm4), edf = df.residual(sc1.fm4))
plot(EMM, comparisons = TRUE)

#emtrends(sc1.fm4, pairwise ~ factor(Group), var = "EYO")
```

```{r}
emmip(sc2.fm1, Group ~ EYO, at=list(EYO=0:15), CIs=TRUE)
Anova(sc2.fm1)

# create a pairwise contrast and export to variable in data.frame form: 
EMM <- emmeans(sc2.fm1, specs=c('Group', 'EYO'), at=list(EYO=eyo_basis))
xprt <- as.data.frame(contrast(EMM, method='pairwise', simple='Group'))
kbl(xprt %>% pivot_wider(id_cols=EYO, names_from=c('contrast'), values_from='p.value'), booktabs = T, caption="p-values (four-way hypothesis testing) for the pairwise comparison between groups at each EYO (LMM - Passive Behavioural phenotype)") %>%
kable_styling(latex_options = c("striped", "scale_down"))

```

### Affective Phenotype
```{r}
eyo_basis <- seq(from=0, to=15, by=1)
N=1e4
# res <- get_preds(sc3.fm6, dataTer.educ, iident, main.effect="Gender", eyo_basis=eyo_basis, N=N)
# kbl(nwayttest(res), booktabs = T, caption="p-values (four-way hypothesis testing) for the pairwise comparison between groups at each EYO (LMM - Affective phenotype)")
plot_effect(sc3.fm6, main.effect="Gender", transform=iident, title='Affective phenotype', ylims=c(-1, 5), eyo_basis=eyo_basis, colors=c('f'="#49b7fc", "m" = "#f38000"))
```

```{r}
emmip(sc3.fm6, EYO ~ Gender)
Anova(sc3.fm6)

#EMM = emmeans(sc1.fm4,  pairwise ~ Gender*Group | EYO)#, at=list(EYO=seq(from=0, to=15, by=1)))
# xprt = print(pairs(EMM), export = TRUE)
# create a pairwise contrast and export to variable in data.frame form: 
EMM <- emmeans(sc3.fm6, specs=c('Gender', 'EYO', 'Education'), at=list(EYO=1:15), nesting = NULL )
xprt <- as.data.frame(contrast(EMM, method='pairwise', simple='Gender'))
kbl(xprt %>% pivot_wider(id_cols=EYO, names_from=c('contrast'), values_from='p.value'), booktabs = T, caption="p-values (four-way hypothesis testing) for the pairwise comparison between groups at each EYO (LMM - Affective phenotype)")%>%
kable_styling(latex_options = c("striped"))

```

#### Contrasts
```{r}
# sc3.fm6 <- lmer(affective ~ 1 + EYO + EYO:Gender + Education + (1|Code), data=dataTer.educ)
joint_tests(sc3.fm6)
emmip(sc3.fm6, Gender ~ EYO, cov.reduce = range)
```

```{r}
# When a covariate and a factor interact, we typically dont want EMMs themselves, but rather estimates of slopes of the covariate trend for each level of the factor. 
trends <- emtrends(sc3.fm6, pairwise ~ Gender:EYO + EYO + Education, var = "EYO")
summary(trends) 

#EMM <- emmeans(sc3.fm6, ".", at=list(Gender=c('m', 'f'), EYO=seq(from=0, to=10)))
#contrast(EMM, "consec", simple = "each", combine = TRUE)
#pairs(EMM, simple = "Gender")    # compare treats for each dose -- "simple effects"
#eff_size(EMM, sigma = sigma(sc3.fm6), edf = df.residual(sc3.fm6))
#plot(EMM, comparisons = TRUE)
```
# Binary Definitive Models

```{r}
library(lme4)
library(car)
library(performance)
library(car)
library(jtools)
# Refit data
# Data ternary
```

```{r}
sc1.bm2 <- lmer(active_behav ~ 1 + EYO + BinGroup + Gender + (1|Code), data=d)
summ(sc1.bm2)
```

```{r}
sc2.bm0 <- lmer(passive_behav ~ 1 + EYO + BinGroup + (1|Code), data=d)
summ(sc2.bm0)
```

```{r}
dred.educ <- d[!is.na(d$Education),]
sc3.bm6 <- lmer(affective ~ 1 + EYO + Gender:EYO + Education + (1|Code) , data=dred.educ)
summ(sc3.bm6)
```

```{r}
export_summs(sc1.bm2, sc2.bm0, sc3.bm6, model.names=c('sc1.bm2', 'sc2.bm0', 'sc3.bm6'),
             error_format = "p={p.value} [{conf.low}, {conf.high}]", number_format = 3)
```

## Figures

```{r}
interact_plot(sc1.bm2, pred=EYO, modx=BinGroup, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Active Behavioural Phenotype', rug=TRUE, jitter=1)
interact_plot(sc2.bm0, pred=EYO, modx=BinGroup, data=d, interval=TRUE, plot.points=FALSE, y.label='Passive Behavioural Phenotype', rug=TRUE, jitter=1)
interact_plot(sc3.fm6, pred=EYO, modx=Education, mod2=Gender, data=d, interval=TRUE, plot.points=FALSE, y.label='Affective Phenotype', rug=TRUE, jitter=1)
```
## Contrasts

```{r}
eyo_basis <- seq(from=0, to=15, by=1)
```

### Active Behavioural Phenotype 



# Bibliography
```{r}
library(report)

cite_packages()
```